---
title: "Feature Selection"
output: html_notebook
---

# Loading Libraries

```{r}
library(Boruta)
library(caret)
library(randomForest)
library(logistf)
library(pROC)  # For AUC
library(caret) # For confusion matrix and other metrics
library(glmnet)
```

# Loading Data

There are two types of data. One contains all the data including demographic information, scale/questionnaire scores and activity level information. Other only contains information about kinematic data for sit to stand motion. The main purpose is to understand the influence of kinematic factors with other factors and as standalone predictors as well.

```{r}
all_data_path <- "../../stats/dataClean2.csv"
kinematic_data_path <- "../../stats/kinematic_features.csv"
all_data <- na.omit(read.csv(all_data_path))
all_data <- all_data[, c(1:ncol(all_data))]
# all_data$fallsBin <- as.factor(all_data$fallsBin)
kinematic_data <- na.omit(read.csv(kinematic_data_path))
kinematic_data <- kinematic_data[, c(1:ncol(kinematic_data))]
summary_all_data <- as.data.frame.matrix(summary(all_data))
summary_kinematic_data <- as.data.frame.matrix(summary(kinematic_data))
head(all_data)
head(kinematic_data)
summary_all_data
summary_kinematic_data
```

# Boruta

Boruta is a randomized wrapper feature selection method which utilizes random forest classification to output variable importance measures.

vs\<-Boruta(class\~features, data=Mydata, pValue = 0.01, mcAdj =TRUE, maxRuns = 100, doTrace=0, getImp = getImpRfZ, ...)

```{r}
set.seed(123)
# data3 <- all_data[, c("fallsBin", predRFE)]
# boruta_all <- Boruta(fallsBin ~ ., data=all_data, pValue=0.05, maxRuns = 1000, doTrace=0)
# boruta_all <- Boruta(fallsBin ~ ., data=data3, pValue=0.05, maxRuns = 1000, doTrace=0)
boruta_kinematic <- Boruta(fallsBin ~ ., data=kinematic_data, pValue=0.05, maxRuns = 1000, doTrace=0)
# print(boruta_all)
# print("----------")
print(boruta_kinematic)
```

```{r}
boruta_signif_all <- getSelectedAttributes(boruta_all, withTentative = TRUE)
boruta_signif_kinematic <- getSelectedAttributes(boruta_kinematic, withTentative = TRUE)
print(boruta_signif_all)
print('-----------')
print(boruta_signif_kinematic)
```

```{r}
# Imp_hist_all<- lapply(
#   1:ncol(boruta_all$ImpHistory),
#   function(i) {
#   boruta_all$ImpHistory[is.finite(boruta_all$ImpHistory[, i]), i]
# })
# # names(Imp_hist_all)<-colnames(boruta_all$ImpHistory)
# # Imp_hist_all <- as.data.frame(Imp_hist_all)
# # Imp_hist_all_selected <- Imp_hist_all[, boruta_signif_all, drop = FALSE]
# # 
# Imp_hist_kinematic<- lapply(
#   1:ncol(boruta_kinematic$ImpHistory),
#   function(i) {
#   boruta_kinematic$ImpHistory[is.finite(boruta_kinematic$ImpHistory[, i]), i]
# })
# names(Imp_hist_kinematic)<-colnames(boruta_kinematic$ImpHistory)
# Imp_hist_kinematic <- as.data.frame(Imp_hist_kinematic)
# Imp_hist_kinematic_selected <- Imp_hist_kinematic[, boruta_signif_kinematic, drop = FALSE]

#summary_lz <- data.frame(unclass(summary(lz)), check.names = FALSE)
#summary_lz <- data.frame(do.call(cbind, lapply(lz, summary)))
#names(lz)<-colnames(boruta_all$ImpHistory)
#lb<-sort(sapply(lz, median))


importance_history <- function(boruta_data) {
  
  boruta_signif <- getSelectedAttributes(boruta_data, withTentative = TRUE)
  
  imp_history <- as.data.frame(boruta_data$ImpHistory)
  
  names(imp_history)<-colnames(boruta_data$ImpHistory)
  
  imp_history_selected <- imp_history[, boruta_signif, drop = FALSE]
  
  # Calculate the median of each column
  col_medians <- apply(imp_history_selected, 2, median)
  
  # Sort the dataframe by column medians (in ascending order)
  imp_history_selected  <- imp_history_selected[, order(col_medians)]
  
  return(imp_history_selected)
}

imp_hist_all <- importance_history(boruta_all)
imp_hist_kinematic <- importance_history(boruta_kinematic)

write.csv(imp_hist_all, "imp_hist_all_falls.csv", row.names = TRUE)
write.csv(imp_hist_kinematic, "imp_hist_kinematic_falls.csv", row.names = TRUE)


```

```{r}
par(family = "Arial", cex.axis = 0.8)
boxplot(imp_hist_all, names = boruta_signif_all, ylab="Importance", las=2)

```

```{r}
imp_all <- boruta_kinematic$ImpHistory[1:14,]

# Calculate the median of each column
col_medians <- apply(imp_all, 2, median)

# Sort the dataframe by column medians (in ascending order)
imp_all  <- imp_all[, order(col_medians)]

write.csv(imp_all, "imp_kinematic.csv", row.names = TRUE)

```

```{r}
shadow_max_all <- as.data.frame(boruta_all$ImpHistory)[, "shadowMax", drop=FALSE]
shadow_min_all <-as.data.frame(boruta_all$ImpHistory)[, "shadowMin", drop=FALSE]
shadow_all <- as.data.frame(c(shadow_max_all, shadow_min_all))
write.csv(shadow_all, "shadow_all.csv", row.names=TRUE)

shadow_max_kinematic <- as.data.frame(boruta_kinematic$ImpHistory)[, "shadowMax", drop=FALSE]
shadow_min_kinematic <-as.data.frame(boruta_kinematic$ImpHistory)[, "shadowMin", drop=FALSE]
shadow_kinematic <- as.data.frame(c(shadow_max_kinematic, shadow_min_kinematic))
write.csv(shadow_kinematic, "shadow_kinematic.csv", row.names=TRUE)
```

```{r}
final_boruta_all<-TentativeRoughFix(boruta_all)
final_boruta_kinematic<-TentativeRoughFix(boruta_kinematic)
print(final_boruta_all)
print(final_boruta_kinematic)
```

```{r}
decision_boruta_all <- as.data.frame(final_boruta_all$finalDecision)
decision_boruta_kinematic <- as.data.frame(final_boruta_kinematic$finalDecision)
```

```{r}
getConfirmedFormula(final_boruta_all)
getConfirmedFormula(final_boruta_kinematic)
```

```{r}
print(as.data.frame(final_boruta_all$finalDecision[final_boruta_all$finalDecision %in% c("Confirmed","Tentative")]))
print(as.data.frame(final_boruta_kinematic$finalDecision[final_boruta_kinematic$finalDecision %in% c("Confirmed","Tentative")]))
```

```{r}
impBoruta_all <- final_boruta_all$finalDecision[final_boruta_all$finalDecision %in%
c("Confirmed")]; length(impBoruta_all)
impBoruta_kinematic <- final_boruta_kinematic$finalDecision[final_boruta_kinematic$finalDecision %in%
c("Confirmed")]; length(impBoruta_kinematic)
```

## Comparison with Recursive Feature Elimination

```{r}
set.seed(123)


features <- all_data[, -ncol(all_data)]  # All columns except the last one
target <- as.factor(all_data[, ncol(all_data)])  # The last column as target

class_counts <- table(target)
class_weights <- c(class_counts[1], class_counts[2])
names(class_weights) <- levels(target)


rfFuncs$fit <- function(x, y, first, last, ...) {
  randomForest(x, y, classwt = class_weights, ...)
}

customSummary <- function(data, lev = NULL, model = NULL) {
  cm <- confusionMatrix(data$obs, data$pred)
  out <- c(specificity = cm$byClass["Specificity"])
  return(out)
}


rfFuncs$summary <- twoClassSummary
control<-rfeControl(functions = rfFuncs, method = "cv", number=10, verbose=TRUE)
rf_train<-rfe(features, target, sizes = c(1:20), rfeControl = control)
rf_train
```

```{r}
plot(rf_train, type=c("g", "o"), cex=1, col=1:5)
```

```{r}
predRFE <- predictors(rf_train)
predBoruta <- getSelectedAttributes(final_boruta_all, withTentative = FALSE)
```

## Comparing with Stepwise Feature Selection

```{ights <- ifelse(y == 1, 371/34, 1)  # Inverse class frequency as weights}


# Fit Elastic Net with weights
cv_fit_weighted <- cv.glmnet(X, y, family = "binomial", alpha = 0.5, weights = class_weights)

# Best lambda from cross-validation
best_lambda_weighted <- cv_fit_weighted$lambda.min

# Final model with best lambda
elastic_net_weighted <- glmnet(X, y, family = "binomial", alpha = 0.5, lambda = best_lambda_weighted, weights = class_weights)

```

```{r}
predictions_prob <- predict(full_model, newx = X, type = "response")

# Convert probabilities to binary class labels (assuming 0.5 threshold)
predictions <- ifelse(predictions_prob > 0.5, 1, 0)

# Convert the actual labels (y) to factor for comparison
actual_labels <- as.factor(y)

# Confusion Matrix
conf_matrix <- confusionMatrix(as.factor(predictions), actual_labels)

# Extract Evaluation Metrics
accuracy <- conf_matrix$overall['Accuracy']
sensitivity <- conf_matrix$byClass['Sensitivity']
specificity <- conf_matrix$byClass['Specificity']
precision <- conf_matrix$byClass['Pos Pred Value']  # Positive Predictive Value
f1_score <- 2 * ((precision * sensitivity) / (precision + sensitivity))

# AUC (Area Under the Curve)
roc_curve <- roc(actual_labels, predictions_prob)
auc <- auc(roc_curve)

# Print the results
cat("Elastic Net Model Evaluation Metrics:\n")
cat("Accuracy:", accuracy, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")
cat("Precision:", precision, "\n")
cat("F1 Score:", f1_score, "\n")
cat("AUC:", auc, "\n")

```
