---
title: "Sit2Stand Analysis"
author: "Melissa Boswell"
editor: "Muhammad Taimoor Adil"
date: "July 27, 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Load libraries and requirements
```{r, echo = FALSE}
library(tidyverse)
library(jtools)
library(FSA)
library(ggpubr)
library(Hmisc)
library(dplyr)
```

# Load data
```{r, echo = FALSE}
df <- read.csv("../../stats/dataClean.csv") # Numeric survey data
```

# Correlation matrix with significance fuction
```{r, echo = FALSE}
# x is a matrix containing the data
# method : correlation method. "pearson"" or "spearman"" is supported
# removeTriangle : remove upper or lower triangle
# results :  if "html" or "latex"
  # the results will be displayed in html or latex format
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```

# Variable name legend
```{r}
df_corr <- df[, c("Age", "Height", "Weight", "BMI", "Sex", "GT", "GPH_TScore", "GMH_TScore", "totalMET", "avgPEM", "avgAAM", "falling_1", "falling_2", "totalSEE", "OA_check", "totalWOMAC", "fallsBin", "falls", "time", "right_ankle_max","right_ankle_min", "right_knee_max","right_knee_min", "right_hip_max","right_hip_min", "trunk_lean_max", "trunk_lean_ang_acc", "pelvic_max_y_acc")]
print("Variable name legend")
print("GPH_TScore: Global physical health t-score")
print("GMH_TScore: Global mental health t-score")
print("OA_check: Whether the individual has osteoarthritis (1) or not (0)")
```
# Boruta df
```{r}
df_boruta_all <- df[, c('ageGroup', "Age", "Height", "Weight", "BMI", "Sex", "fallsBin", "numMedCond", "pelvic_avg_speed", "right_hip_min_ang_acc", "right_hip_max_mean_stand2sit", "left_hip_min_ang_vel_sit2stand", "right_ankle_max_sd_stand2sit", "left_knee_max_mean_stand2sit")]
df_boruta_kin <- df[, c('ageGroup', "Age", "Height", "Weight", "BMI", "Sex", "fallsBin", 'left_hip_min_ang_vel_sit2stand', 'left_knee_min_ang_vel_stand2sit', 'neck_min_y_acc', 'pelvic_avg_speed', 'right_ankle_max_sd_stand2sit', 'right_hip_max_mean', 'right_hip_max_mean_stand2sit', 'right_hip_max_stand2sit', 'right_hip_min_ang_acc', 'speed', 'speed_stand2sit', 'time_diff_sit2stand', 'time_diff_stand2sit', 'time_stand2sit')]
```

# Elastic Net -> Selection Frequency df
```{r}
df_elastic_all <- df[, c('ageGroup', "Age", "Height", "Weight", "BMI", "Sex", "fallsBin", 'falling_1', 'speed_sd', 'right_ankle_ang_acc','right_ankle_ang_vel', 'vigMins', 'walkMET', 'walkTime', 'Meditation', 'trunk_lean_min_ang_acc', 'numMedCond')]
df_elastic_kin <- df[, c('ageGroup', "Age", "Height", "Weight", "BMI", "Sex", "fallsBin", 'speed_sd', 'right_hip_max', 'right_ankle_ang_acc','right_ankle_ang_vel', 'left_shank_angle_ang_acc', 'left_shank_angle_ang_vel', 'left_shank_angle_min_sd_stand2sit', 'pelvic_avg_y_acc_stand2sit', 'pelvic_avg_y_acc', 'pelvic_avg_y_acc_sit2stand')]
```

# Elastic Net -> Coefficient x Frequency df
```{r}
df_csf_all <- df[, c('ageGroup', "Age", "Height", "Weight", "BMI", "Sex", "fallsBin", 'speed_sd', 'IPAQ_SF_5', 'pelvic_avg_y_acc', 'pelvic_avg_y_acc_sit2stand', 'pelvic_avg_y_acc_stand2sit', 'numMed_2', 'pelvic_avg_y_speed_stand2sit', 'pelvic_avg_y_speed', 'pelvic_avg_y_speed_sit2stand', 'falling_2')]
df_csf_kin <- df[, c('ageGroup', "Age", "Height", "Weight", "BMI", "Sex", "fallsBin", 'speed_sd', 'pelvic_min_speed_stand2sit', 'pelvic_avg_y_acc', 'pelvic_avg_y_acc_sit2stand', 'pelvic_avg_y_acc_stand2sit','neck_avg_y_acc', 'pelvic_min_speed_sit2stand','pelvic_avg_y_speed_stand2sit', 'pelvic_avg_y_speed_sit2stand','pelvic_avg_y_speed')]
```

# Selected All without correlations
```{r}
df_selected <- df[, c('left_knee_max_mean_stand2sit', 'neck_min_y_acc',
       'pelvic_avg_speed', 'trunk_lean_min_ang_acc',
       'right_ankle_max_sd_stand2sit', 'pelvic_min_speed_stand2sit',
       'right_hip_max_mean',
       'left_knee_min_ang_vel_stand2sit',
       'falling_1', 'time_stand2sit', 'Meditation',
       'left_hip_min_ang_vel_sit2stand',
       'right_ankle_max_mean_stand2sit', 'falling_2',
       'pelvic_avg_y_speed',
       'left_shank_angle_min_sd_stand2sit',
       'right_hip_max_mean_stand2sit',
       'right_ankle_ang_acc', 'pelvic_avg_y_acc',
       'vigMins', 'right_hip_max',
       'right_hip_min_ang_acc',
       'numMedCond',
       'left_shank_angle_ang_acc', 'neck_avg_y_acc',
       'left_hip_min_ang_vel_sit2stand',
       'speed_sd', 'walkTime',
       'right_hip_min_ang_acc',
       'numMed_2', 'IPAQ_SF_5',
       'pelvic_min_speed_sit2stand')]

df_selected_final <- df[, c('fallsBin','left_knee_max_mean_stand2sit', 'neck_min_y_acc',
       'pelvic_avg_speed', 'trunk_lean_min_ang_acc',
       'right_ankle_max_sd_stand2sit', 'left_knee_min_ang_vel_stand2sit',
       'falling_1', 'time_stand2sit', 'Meditation',
       'left_hip_min_ang_vel_sit2stand', 'right_ankle_max_mean_stand2sit',
       'falling_2', 'pelvic_avg_y_speed',
       'left_shank_angle_min_sd_stand2sit', 'right_ankle_ang_acc',
       'pelvic_avg_y_acc', 'vigMins', 'right_hip_max',
       'right_hip_min_ang_acc', 'numMedCond', 'left_shank_angle_ang_acc',
       'neck_avg_y_acc', 'speed_sd', 'walkTime', 'IPAQ_SF_5',
       'pelvic_min_speed_sit2stand')]


```

# Filter age groups 50 plus and 50 minus
```{r}
df_age50plus <- filter(df, ageGroup == "[50-80+")
df_age50minus <- filter(df, ageGroup != "[50-80+")
df_boruta_all_50plus <- filter(df_boruta_all, ageGroup == "[50-80+")
df_boruta_kin_50plus <- filter(df_boruta_kin, ageGroup == "[50-80+")
df_elastic_all_50plus <- filter(df_elastic_all, ageGroup == "[50-80+")
df_elastic_kin_50plus <- filter(df_elastic_kin, ageGroup == "[50-80+")
df_csf_all_50plus <- filter(df_csf_all, ageGroup == "[50-80+")
df_csf_kin_50plus <- filter(df_csf_kin, ageGroup == "[50-80+")
```

# Descriptive tables
```{r}
# Removed subjects
#df_removed <- df_all[! df_all_text$subjectid %in% df_text$subjectid,]

# Simple data frames
df.descr <- df[, c("Age", "Sex", "Height", "Weight", "BMI", "OA_check", "GPH_TScore", "GMH_TScore")]
df_age50plus.descr <-df_age50plus[, c("Age", "Sex", "Height", "Weight", "BMI", "OA_check", "GPH_TScore", "GMH_TScore")]

describe(df.descr)
describe(df_age50plus.descr)
```

# View age bins
```{r}
summary(df$ageGroup)
```

# Overall timing
# Mean Time
```{r}
# 18+
summary(df$time_sit2stand)
# Standard deviation
sd(df$time_sit2stand)
# Age bins
tapply(df$time_sit2stand, df$ageGroup, summary)
# Standard deviation
tapply(df$time_sit2stand, df$ageGroup, sd)

# T-tests
x <- t.test(df$time_sit2stand[row(df)[which(df$ageGroup == "18-50)")]], df$time_sit2stand[row(df)[which(df$ageGroup == "[50-80+")]])
x
x$conf.int

# Box Plot
TimeAgeGroup <- tapply(df$time_sit2stand, df$ageGroup, summary)
boxplot(TimeAgeGroup, xlab="Age Group", ylab="Time")
```
```{r}
sig_test <- function(split_group, split_values, check_parameter, test){
  # 18+
  summary(check_parameter)
  # Standard deviation
  sd(check_parameter)
  # Bins
  tapply(check_parameter, split_group, summary)
  # Standard deviation
  tapply(check_parameter, split_group, sd)
  
  if(test == "t"){
  # T-tests
  x <- t.test(check_parameter[row(df)[which(split_group == split_values[1])]], check_parameter[row(df)[which(split_group == split_values[2])]])
  print(x)
  print(x$conf.int)
  }
  else if(test == "w"){
  # Mann-Whitney U test (Wilcoxon Rank Sum)
  x <- wilcox.test(check_parameter[row(df)[which(split_group == split_values[1])]], check_parameter[row(df)[which(split_group == split_values[2])]])
  print(x)
  print(x$conf.int)
  }
  
  # Box Plot
  TimeAgeGroup <- tapply(check_parameter, split_group, summary)
  boxplot(TimeAgeGroup, xlab="Age Group", ylab="Time")
}

sig_test(df$ageGroup, c("18-50)", "[50-80+"), df$fallsBin, "w")


```

# Correlations

```{r}
corstars(df[, c("time", "trunk_lean_max_mean_sit2stand", "trunk_lean_max_ang_acc_sit2stand", "Age", "Sex", "BMI", "EthBin", "GPH_TScore", "GMH_TScore", "OA_check")])
corstars(df_age50plus[, c("time", "trunk_lean_max_mean_sit2stand", "trunk_lean_max_ang_acc_sit2stand", "GPH_TScore", "GMH_TScore")])
```

```{r}
columns_to_exclude <- c('ageGroup', "Age", "Height", "Weight", "BMI", "Sex", "fallsBin")

df_clean <- df_csf_all %>%
  select(-all_of(columns_to_exclude))

corstars(df_selected_final)
#write.csv(corstars(df_selected), "cor_selected.csv", row.names = T)
```


# Logistic Regression Models
```{r, echo = FALSE}
# Binary Fall status
# Selected parameters status - control for time, age, sex, BMI

columns_to_exclude <- c("ageGroup", "Height", "Sex", "Age", "BMI", "Weight")

df_clean <- df_csf_all_50plus %>%
  select(-all_of(columns_to_exclude))


# Scale only continuous (numeric) variables and keep categorical variables unchanged
df_scaled <- df_selected_final %>%
  mutate_if(is.numeric, scale)

mod <- glm(fallsBin ~ ., data = df_scaled)
summary(mod, family = "binomial")
confint(mod)
exp(coef(mod))
```

# Analyses
```{r, echo = FALSE}
# Time predict GPH - control for age, sex, BMI
summ(lm(df$GPH_TScore ~ scale(df$time) + scale(df$Age) + df$Sex + scale(df$BMI)),confint = TRUE, digits = 3)

# Trunk lean max predict ethnicity status - control for time, age, sex, BMI
# Kruskal-Wallis test
## Because one ethnic group not normally distributed
df.eth <- rbind(cbind(df$trunk_lean_max_mean_sit2stand[which(df$Ethnicity==1)], df$time[which(df$Ethnicity==1)], "White"), cbind(df$trunk_lean_max_mean_sit2stand[which(df$Ethnicity==2)], df$time[which(df$Ethnicity==2)], "Black"),  cbind(df$trunk_lean_max_mean_sit2stand[which(df$Ethnicity==3)], df$time[which(df$Ethnicity==3)], "Hispanic"), cbind(df$trunk_lean_max_mean_sit2stand[which(df$Ethnicity==4)], df$time[which(df$Ethnicity==4)], "Asian"))
df.eth <- data.frame(df.eth)
colnames(df.eth) <- c("trunk.lean", "time", "ethnicity")
df.eth$ethnicity <- as.factor(df.eth$ethnicity)
df.eth$trunk.lean <- as.numeric(df.eth$trunk.lean)
df.eth$time <- as.numeric(df.eth$time)

## Compute summary statistics
group_by(df.eth, ethnicity) %>%
  summarise(
    count = n(),
    mean = mean(trunk.lean, na.rm = TRUE),
    sd = sd(trunk.lean, na.rm = TRUE),
    median = median(trunk.lean, na.rm = TRUE),
    IQR = IQR(trunk.lean, na.rm = TRUE)
  )

## Kruskal-Wallis
kruskal.test(trunk.lean ~ ethnicity, data = df.eth)

## Dunn's Test with Bonferroni correction for p-values
dunnTest(trunk.lean ~ ethnicity, data = df.eth, method = "bh")

# Control for other variables
df.eth.bin <- data.frame(rbind(cbind(df$trunk_lean_max_mean_sit2stand[which(df$Ethnicity==1)], df$time[which(df$Ethnicity==1)], df$Age[which(df$Ethnicity==1)], df$Sex[which(df$Ethnicity==1)], df$BMI[which(df$Ethnicity==1)], df$GPH_TScore[which(df$Ethnicity==1)], 0), cbind(df$trunk_lean_max_mean_sit2stand[which(df$Ethnicity==4)], df$time[which(df$Ethnicity==4)], df$Age[which(df$Ethnicity==4)], df$Sex[which(df$Ethnicity==4)], df$BMI[which(df$Ethnicity==4)], df$GPH_TScore[which(df$Ethnicity==4)], 1)))
colnames(df.eth.bin) <- c("trunk.lean", "time", "Age", "Sex", "BMI", "health", "ethnicity")

## Binned ethnicity
mod2 <- glm(formula = df.eth.bin$ethnicity ~ scale(df.eth.bin$trunk.lean) + scale(df.eth.bin$time) + scale(df.eth.bin$Age) + df.eth.bin$Sex + scale(df.eth.bin$BMI) + scale(df.eth.bin$health))

summary(mod2, family = "binomial")
confint(mod2)
exp(coef(mod2))

# OA status
# Trunk lean max predict OA status - control for time, age, sex, BMI

mod3 <- glm(df$OA_check ~ scale(df$time) + scale(df$Age) + df$Sex + scale(df$BMI))

summary(mod3, family = "binomial")
confint(mod3)
exp(coef(mod3))

mod4 <- glm(df$OA_check ~ scale(df$trunk_lean_max_mean_sit2stand) + scale(df$time) + scale(df$Age) + df$Sex + scale(df$BMI))

summary(mod4, family = "binomial")
confint(mod4)
exp(coef(mod4))

# Global mental health
# Trunk acceleration max predict GMH - control for time, age, sex, BMI
mod4 <- lm(df_age50plus$GMH_TScore ~ scale(df_age50plus$trunk_lean_max_ang_acc_sit2stand) + scale(df_age50plus$time) + scale(df_age50plus$Age) + df_age50plus$Sex + scale(df_age50plus$BMI))

summ(mod4,confint = TRUE, digits = 3)
confint(mod4)

```

# Plots
```{r, echo = FALSE}
# Age Bins
TimeAgeGroup <- tapply(df$time, df$ageBin, summary)
boxplot(TimeAgeGroup, xlab="Age Bin", ylab="Time")

# Physical health vs. Time
plot1 <- ggscatter(df_age50plus, x = "right_hip_max_mean_sit2stand", y = "Age" , xlab = "Age", size = 3, font.label = c(20, "plain"),
  ylab = "right_hip_max_mean",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "lightblue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
plot1 + stat_cor(method = "pearson", label.x = 10, label.y = 70)

# Age vs. Trunk Lean
plot2 <- ggscatter(df_age50plus, x = "trunk_lean_max_mean_sit2stand", y = "Age", xlab = "Max Trunk Lean", size = 3, font.label = c(20, "plain"),
  ylab = "Age",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "lightblue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
plot2 + stat_cor(method = "pearson", label.x = 190, label.y = 90)

# Ethnicity vs. Trunk lean
TrunkLean_Eth <- tapply(df$trunk_lean_max_mean_sit2stand, df$Ethnicity, summary)
boxplot(TrunkLean_Eth, xlab="Ethnicity", ylab="Max Trunk Lean")

# Osteoarthritis Status vs. trunk angle
df$OA_check.fact <- as.factor(df$OA_check)
ggplot(df, aes(x=OA_check.fact, y=trunk_lean_max_mean_sit2stand)) + 
  geom_boxplot(fill="gray")+
  labs(x="Osteoarthritis Status", y = "Max Trunk Lean")+
  theme_classic()

x <- t.test(df$trunk_lean_max_mean_sit2stand[row(df)[which(df$OA_check == 0)]], df$trunk_lean_max_mean_sit2stand[row(df)[which(df$OA_check == 1)]])
x
x$conf.int

# Mental health vs. Trunk Acceleration
plot3 <- ggscatter(df_age50plus, x = "trunk_lean_max_ang_acc_sit2stand", y = "GMH_TScore", xlab = "Max Trunk Acceleration", size = 3, font.label = c(20, "plain"),
  ylab = "Mental Health",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "lightblue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
plot3 + stat_cor(method = "pearson", label.x = 250, label.y = 70)
```

# Corrections for multiple comparisons
```{r, echo = TRUE}
all.cors.p <- data.frame(cor.test(df$time,df$Age)$p.value, cor.test(df$time,df$Sex)$p.value, cor.test(df$time,df$BMI)$p.value, cor.test(df$time,df$EthBin)$p.value, cor.test(df$time,df$GPH_TScore)$p.value, cor.test(df$time,df$GMH_TScore)$p.value, cor.test(df$time,df$OA_check)$p.value, cor.test(df$trunk_lean_max_mean_sit2stand,df$Age)$p.value, cor.test(df$trunk_lean_max_mean_sit2stand,df$Sex)$p.value, cor.test(df$trunk_lean_max_mean_sit2stand,df$EthBin)$p.value, cor.test(df$trunk_lean_max_mean_sit2stand,df$BMI)$p.value, cor.test(df$trunk_lean_max_mean_sit2stand,df$GPH_TScore)$p.value, cor.test(df$trunk_lean_max_mean_sit2stand,df$GMH_TScore)$p.value, cor.test(df$trunk_lean_max_mean_sit2stand,df$OA_check)$p.value, cor.test(df$trunk_lean_max_ang_acc_sit2stand,df$Age)$p.value, cor.test(df$trunk_lean_max_ang_acc_sit2stand,df$Sex)$p.value, cor.test(df$trunk_lean_max_ang_acc_sit2stand,df$BMI)$p.value, cor.test(df$trunk_lean_max_ang_acc_sit2stand,df$EthBin)$p.value,  cor.test(df$trunk_lean_max_ang_acc_sit2stand,df$GPH_TScore)$p.value, cor.test(df$trunk_lean_max_ang_acc_sit2stand,df$GMH_TScore)$p.value, cor.test(df$trunk_lean_max_ang_acc_sit2stand,df$OA_check)$p.value)

all.cors.p.adj <- p.adjust(all.cors.p, method = "fdr", n = length(all.cors.p))
all.cors.p.adj

all.cors.50.p <- data.frame(cor.test(df_age50plus$time,df_age50plus$GPH_TScore)$p.value, cor.test(df_age50plus$time,df_age50plus$GMH_TScore)$p.value, cor.test(df_age50plus$trunk_lean_max_mean_sit2stand,df_age50plus$GPH_TScore)$p.value, cor.test(df_age50plus$trunk_lean_max_mean_sit2stand,df_age50plus$GMH_TScore)$p.value, cor.test(df_age50plus$trunk_lean_max_ang_acc_sit2stand,df_age50plus$GPH_TScore)$p.value, cor.test(df_age50plus$trunk_lean_max_ang_acc_sit2stand,df_age50plus$GMH_TScore)$p.value)

all.cors.50.p.adj <- p.adjust(all.cors.50.p, method = "fdr", n = length(all.cors.50.p))
all.cors.50.p.adj
```